---
title: "Libraries as Urban Infrastructure"
output: html_document
---

```{r, include = FALSE}
source("R/01_helper_functions.R")
options(cancensus.api_key = "CensusMapper_4be2cf3bf91d0cabf967f4934dbdc63b")
source("R/02_data_import.R")
source("R/03_service_area_comparison.R")
source("R/06_final_figures")
```

As libraries take on a greater role as community centres and fill social-service gaps, we need to understand the challenges they confront as a result of changing demographics and social needs in their local areas.

This preliminary study was requested by the Municipal Council of Urban Libraries. We conducted a spatial analysis, mapping over 500 urban libraries and a series of demographic variables to determine how demographics differed within library service areas and the wider city context.
Overall, findings demonstrate systematically higher levels of social need within library walk-sheds, with higher core housing need, lower income, higher unemployment, and more visible minorities and immigrants when compared with surrounding areas.

```{r, echo = FALSE, fig.cap="\\label{figure1} Figure 1: Distribution of housing need in Motnreal", fig.align = 'center', message = FALSE}
figure_3
```
*Core housing need is 57% higher within library walksheds in Montreal than in the rest of the CMA.*

Comparing the census data from 2006 and 2016, we found that these patterns of greater social need in library walksheds are not new; social disparities between library service areas and the rest of the regions are relatively large and mostly growing. 

```{r, echo = FALSE, fig.cap="\\label{figure1} Figure 2: Housing income and library walksheds", fig.align = 'center', message = FALSE}
figure_5b
```


Changing patterns of urban development have shifted, and will continue to shift, libraries’ local contexts and roles. With growing social need in library walk-sheds, libraries are becoming increasingly important spaces for community services, moving away from the traditional mission of public libraries. However, evidence suggests that while municipal officials are supportive of libraries, they do not fully understand the range of services that they may provide; stagnating or declining high-level funding is a major threat.



